Markov Link Method for calibrating without joint measurement, including the case of destructive measurements

### 2018-07-19 paper/paper.tex
Changed
Latest and greatest

\newcommand{\figgygr}[2]{\includegraphics[width=#1\textwidth]{#2}}
% \newcommand{\figgygr}[2]{picture here}
\newcommand{\figgy}[1]{\begin{figure}\fbox{\begin{minipage}{\textwidth}#1\end{minipage}}\end{figure}}

@@

Formally, a calibration is simply a conditional distribution.  We will denote it by $q^*(y|x)$.  Let us say we obtain measurement result $x$ from one tool on a particular specimen.  The number $q(y|x)$ indicates the probability of obtaining result $y$ from a second tool applied to measure the same specimen.  One way to learn the calibration is to measure the same specimens with both tools.  We call this ``joint measurement.''  Unfortunately, calibrations are often required even when joint measurement is unavailable.  For example, if the measurement tool significantly alters the specimen being measured, joint measurement is simply impossible.  In other cases, it may be expensive or impractical.

We here propose the Markov Link Method (MLM) to estimate calibrations between tools.  The MLM can be applied without any joint measurement.  The key idea is to use multiple subpopulations of specimens.  If each subpopulation captures a different slice of the overall population, we can obtain tight bounds on the true calibration.  This is true even if each subpopulation is highly heterogeneous.  By integrating information from all the subpopulations we can make rigorous deductions about what the calibration might be.  MLM also gives suggestions about which further subpopulations might be helpful to study in order to further refine our knowledge of the true calibration.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Markov Link Method assumption}

The MLM is only applicable if a certain assumption is met.  In this section we will articulate this assumption as clearly as possible.  Let us say we are considering the specimens of a large population.  For example, each specimen in the large population might be a human cell.  Each specimen could also be be a piece of metal which needs to be tested.  We will assume there are three basic properties of interest for each specimen $i$:

\begin{enumerate}
\item $\ell_i$, the subpopulation.  We assume that we can identify a variety of these subpopulations.  They are not required to be disjoint subpopulations; it is only necessary that each subpopulation captures a different slice of the overall population.  For example, we could define subpopulations by looking at cells in different parts of the body or cells with different sizes.  
\item $X_i$, the result of measuring a specimen with tool I.  For example, perhaps tool I takes a picture of the cell with a destructive electron microscopy method.
\item $Y_i$, the result of measuring a specimen with tool II.  For example, perhaps tool II measures the RNA expression of the cell with a destructive sequencing method.
\end{enumerate}

We here consider the case that ``joint measurement'' is impossible or impractical.  In terms of the notation above, that means that for any given specimen we can either observe $\ell_i, X_i$ or $\ell_i,Y_i$.  We can never observe $\ell_i,X_i,Y_i$ for any specimen $i$.  We would like to estimate as much as possible about the distribution of $X,Y|\ell$.  In this paper, we will consider the case that this estimation is facilitated by an additional assumption:

\figgy{
\hfill{}\figgygr{.7}{../images/bigpicture}\hfill{}
\caption{We consider the case that for each specimen there are two ways we might measure it.  We divide the specimens into two groups.  The specimens on the left are measured with technique I and the specimens on the right are measured with technique II.  In the end, we are left wondering: what would have happened if we had measured the specimens on the left with technique II?  Or what if we had measured the specimens on the right with technique I?  The Markov Link Method gives a way to answer this problem using a subpopulation analysis.   \label{fig:bigpicture}}
}

\begin{center}
\fbox{\parbox[c]{4in}{
\vspace{.1in}
\begin{center}
\textbf{The Markov Link Method Assumption}\\$\mathbb{P}(Y=y|X=x,\ell)=q^*(y|x)$ for every value of $\ell$.
\end{center}
\vspace{.1in}
}}
\end{center}

Intuitively, this signifies that the manner in which $X$ predicts $Y$ would be the same for each subpopulation.  In the language of statistics, the assumption is that the variable $X$ is `sufficient' for $\ell$.  If this assumption is not met, then the method presented here is not applicable.  If it \emph{is} met, it opens the door for new methods to estimate three important objects:

\begin{itemize} 
    \item The calibration $q^*(y|x)$ 
    \item The distribution of tool-I measurement values found in each subpopulation, $p^*(x|\ell)$ 
    \item The distribution of tool-II measurement values found in each subpopulation,  $h^*(y|\ell)$.  
\end{itemize}

Under the MLM assumption, it is straightforward to see that these three objects are all bound together by the formula $h^*(y|\ell) = \sum_x p^*(x|\ell)q^*(y|x)$ for each $\ell,y$.  Using this equation, we will see that it is possible to estimate properties of the calibration $q^*$ even without access to joint measurement.  It also makes it possible to perform more efficient estimation of $h^*$.  Theoretically we could also get more efficient estimation of $p^*$, but we did not find plausible cases where this actually occured.

\figgy{
\hfill{}\figgygr{.7}{../images/plate}\hfill{}
\caption{Another way to articulate the Markov Link Method is to look at the graphical model associated with the MLM assumption.  This reveals that we have two kinds of data.  In one case (above), we can observe $\ell,X$, but $Y$ is hidden.  For other kinds of data (below) we can observe $\ell,Y$, but $X$ is hidden.  All of this data is joined by the unknown parameters $p,q$.  Here $p$ indicates the distribution of tool-I measurement values from each subpopulation, and $q$ indicates the calibration.  Our task is to use these different kinds of data to determine the parameters $p,q$ as best as we can.  Once these parameters are known, notice that distribution of tool-II measurement values for each subpopulation, which we denote $h$, can be determined immediately using the Markov Link Assumption: $h^*(y|\ell) = \sum_x p^*(x|\ell)q^*(y|x)$.  \label{fig:platediag}}
}

@@

@@

\begin{itemize}
    \item Quality control for manufacturing.   One way to test the reliability of a part is to construct a machine that pushes the part until it breaks.  However, how can we test the reliability of the machine that performs the test?  In each test run there will be some variability induced by the machine itself, which induces a measurement error.  In practice, some kind of assumptions about part homogeneity are used to approximate this error (cf.\ \cite{de2005gauge}).  However, if we have two testing machines we can use the MLM to obtain a calibration between the machines, even though we can never test the same part with both machines.  This enables us to bound the overall measurement error.  In this case, $\ell$ might indicate the type of a part being tested, $X$ would indicate the reliability of a part as measured by one machine, and $Y$ would indicate the reliability of a part as measured by another machine.  If the error in machine $Y$ is not correlated to the part type $\ell$, then the MLM assumption certainly holds.  Even if the error is correlated, the MLM assumption may still hold.  For example, imagine that the $Y$ error is correlated with the absolute reliability of the part, this may pose no problem if that reliability is adequately measured by $X$.

    \item Combining knowledge across experimental modalities: morphology and transcriptomics.  There are different ways to think about the different types of cells in an organism.  A traditional approach is to classify cells based on what they look like (cf.\ \cite{steinman1973identification,bloomfield1982physiological}).  A more modern approach is to assay the cell's transcriptome (cf.\ \citep{tasic2017shared}).  Unfortunately, modern high-resolution cell photography and single-cell sequencing technologies are both destructive.  As a result, we can't always get both kinds of data for the same specimens.  For cells native to regions full of diverse cell-types, it is thus quite hard to grasp the correspondence between these different kinds of classification systems.  The result is two completely independent classifications of cells, one for each way of looking at the cell.  MLM allows us to estimate the relationship between those two classification systems, yielding a wholistic understanding of the different types of cells.  In this case, $\ell$ might indicate some side information such as where in the body the cell was found, $X$ would indicate a detailed classification of the cell according to its transcriptomics, and $Y$ would indicate a coarser classification of a cell according to its morphology.  We expect that cell morphology is largely a function of cell transcriptomics.  Thus, as long as the $X$ measurement is sufficiently detailed, we expect that any correlations between $Y$ and $\ell$ would be explained by $X$.  That is, the MLM assumption holds.

@@

    \item Text/image correspondence.  Automatic image captioning is an ongoing effort in machine learning (cf.\ \cite{srivastava2018survey}).  There are three types of data available to help develop such algorithms: text-only data, image-only data, and paired-text-and-image data.  Obviously the last kind is the most useful for automatic image captioning, but there is much less of it.  The Markov Link Method suggests one way to use the more plentiful text-only and image-only data.  We can first apply classic machine learning techniques to get coarse labels for both kinds of data.  Using this side-information to identify subpopulations, the MLM can then deduce a fine-grained correspondence between text and images by combining information from across all the subpopulations.  Here $\ell$ would indicate coarse labels such as ``cat'' or ``street scene.''  These labels could be derived from either images or text and can be trained in a supervised fashion.  $X$ would indicate the image and $Y$ would indicate a caption.  If caption is largely determined by the picture $X$, the MLM assumption may hold.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Markov Link Method: input and output}

The Markov Link Method is a procedure from estimating the distribution $X,Y|\ell$ from data, under the Markov Link Method assumption.  Details on the exact algorithm and convergence results are given in Section \ref{sec:mlm}.  For now, let us content ourselves with looking at the input and the output of the MLM:

\begin{description}
    \item[Input] We assume we are given $n$ samples using tool I and $m$ samples using tool II.
    \begin{itemize}
        \item Samples $(\ell_1,X_1)\cdots (\ell_n,X_n)$ sampled such that $\mathbb{P}(X_i=x|\ell_i=\ell)= p^*(x|\ell)$
        \item Samples $(\ell_{n+1},Y_{n+1})\cdots (\ell_{n+m},Y_{n+m})$ sampled such that $\mathbb{P}(Y_i=x|\ell_i=\ell)= h^*(x|\ell)$
    \end{itemize}
    We assume that $\ell,X,Y$ are discrete random variables with finite support. We also assume that the MLM assumption holds, i.e.\ $h^*(x|\ell)=\sum_y p^*(x|\ell)q^*(x|y)$ for some unknown $q^*$.  It would be interesting to generalize the method to non-discrete random variables, but we leave this for future work.  Note that if the random variables are low-dimensional and continuous, the MLM can still be applied by discretizing the continuous space into a finite number of bins.
    \item[Output] Our goal is to estimate $p^*,q^*,h^*$ from the samples.  In this effort, we return:
    \begin{itemize}
        \item Point estimates, $\hat p,\hat q,\hat h$.  
        \item Quantitative estimates of the variability of these point estimates due to insufficient data, $\Delta_p,\Delta_q,\Delta_h$.  
        \item An quantitative estimate of the variability of $\hat q$ due to so-called `identifiability' issues, $\bar d$.  
        \item Variability visualizations indicating a set of ``plausible'' values for each row of $q$; this gives a qualitative counterpart to the quantitative measures of variability.
    \end{itemize}
    Precise definitions and mathematical results are given in section \ref{sec:mlm}.
\end{description}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulation results}

Before we show the results of the MLM on real data, let us look at some simulation results to see what the MLM can and cannot achieve.  In particular, we will:

\begin{itemize}
    \item Pick a set of possible parameter values for $p^*,q^*,h^*$.
    \item Generate data according to these distributions, namely $n$ samples according to $p^*$ and $m$ samples according to $h^*$.  
    \item Pretend to forget the true values of $p^*,q^*,h^*$.
    \item Apply the Markov Link Method to the generated samples.
    \item Compare the results of the MLM with the original true values.
\end{itemize}

Our purpose is to see how well the method performs if all of its assumptions are met. We will measure performance two ways:

\begin{enumerate}
    \item First, we will measure the error of the estimators $\hat p,\hat q,\hat h$, i.e.\ how much they differ from the true generating distributions $p^*,q^*,h^*$.  For $h$ we will compare this measurement error against a standard pseudocount estimator (we will not do the same for $p$, because the MLM estimator for $p$ \emph{is} a pseudocount estimator).  We refer the reader to Section \ref{sec:mlm} for details on this estimator.
    \item Second, we will look at how well the MLM has estimated its uncertainty about the estimates.  We know that there will always be some amount of error.  For this reason, the MLM offers two tools to estimate the amount of error that may be present.  In this section we will particularly focus on the quantities $\Delta_p,\Delta_q,\Delta_h$ (which approximately indicate variability due to insufficient data) and $\bar d$ (which approximately indicates variability due to so-called identifiability concerns).  For a precise explanation of these two sources of variability, we refer the reader to Section \ref{sec:mlm}.  
\end{enumerate}

>>>>

\newcommand{\figgygr}[2]{\includegraphics[width=#1\textwidth]{#2}}
\newcommand{\figgygrs}[3]{\begin{tabular}{c}\includegraphics[width=#1\textwidth]{#2}\\#3\end{tabular}}
\newcommand{\figgygrsTable}[3]{\begin{tabular}{c}{#2}\\#3\end{tabular}}
\newcommand{\figgygrsH}[3]{\begin{tabular}{c}\includegraphics[width=#1\textwidth,angle=270]{#2}\\#3\end{tabular}}

% \newcommand{\figgygr}[2]{pic}
% \newcommand{\figgygrs}[3]{\begin{tabular}{c}pic\\#3\end{tabular}}
% \newcommand{\figgygrsTable}[3]{\begin{tabular}{c}{#2}\\#3\end{tabular}}
% \newcommand{\figgygrsH}[3]{\begin{tabular}{c}pic\\#3\end{tabular}}

\newcommand{\figgy}[1]{\begin{figure}\fbox{\begin{minipage}{\textwidth}#1\end{minipage}}\end{figure}}

@@

Formally, a calibration is simply a conditional distribution.  We will denote it by $q^*(y|x)$.  Let us say we obtain measurement result $x$ from one tool on a particular specimen.  The number $q^*(y|x)$ indicates the probability of obtaining result $y$ from a second tool applied to measure the same specimen.  One way to learn the calibration is to measure the same specimens with both tools.  We call this ``joint measurement.''  Unfortunately, calibrations are often required even when joint measurement is unavailable.  For example, if the measurement tool significantly alters the specimen being measured, joint measurement is simply impossible.  In other cases, it may be expensive or impractical.

We here propose the Markov Link Method (MLM) to estimate calibrations between tools.  The MLM can be applied without any joint measurement.  The key idea is to use multiple subpopulations of specimens.  If each subpopulation captures a different slice of the overall population, we can obtain tight bounds on the true calibration.  This is true even if the subpopulations are highly heterogeneous and overlapping.  By integrating information from all the subpopulations we can make rigorous deductions about what the calibration might be.  MLM also gives suggestions about which further subpopulations might be helpful to study in order to further refine our knowledge of the true calibration.

\figgy{
\hfill{}\figgygrs{.5}{../images/bigpicture}{(a) Setup: each specimen can be \\ measured with either tool I or tool II, \\ but can't be measured with both.\\ How do we calibrate the tools?}
\hfill{}\figgygrs{.4}{../images/plate}{(b) Key modeling assumption: \\ the calibration $q$ is the same \\ for every subpopulation. }\hfill{}

\caption{{\bf The Markov Link Method assumption makes calibration possible, even without joint measurement}.  In subfigure (a) we measure some neurons with tool I and others with tool II.  How can we calibrate the tools?  That is, how can we know what we might have seen if we had measured the first group with tool II or the second group with tool I?  The Markov Link Method (MLM) gives a way to get a partial answer, using a subpopulation analysis.  In subfigure (b) we present a plate diagram describing our beliefs about the conditional independencies in this situation; cf.\ \cite{koller2009probabilistic} for a thorough explanation of this type of diagram.  For each specimen we are interested in $\ell$ (subpopulation), $X$ (tool-I measurement), and $Y$ (tool-II measurement).  For $n$ samples we can see $(\ell,X)$ and for $m$ samples we can see $(\ell,Y)$.  However, for the first group of $n$ samples we cannot observe $Y$, and for the second group of $m$ samples we canot observe $X$.  This is indicated in the diagram by the grayed-out circles.   This lack of joint measurement makes it hard to learn the variable distributions.  The MLM has a key assumption that makes it possible: we assume $X$ is sufficiently detailed to \emph{statistically isolate} $Y$ from $\ell$.  This assumption is reflected in the arrows of the diagram; all paths from $\ell$ variables to $Y$ variables pass through $X$ variables.  The assumption allows us to learn much about the important probability distributions, as parametrized by the tool-I measurement distribution $p(x|\ell)$ and the calibration $q(y|x)$. \label{fig:bigpicture}}
}

This paper will proceed in the following sections:
\begin{itemize}
    \item Description of how to apply the Markov Link Method, using an example to explain the process
    \item A variety of examples where the method may apply
    \item The method's performance in simulation
    \item The method's results on real-world data
    \item Relation to prior work
    \item Discussion
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%            _           
%  _ __ ___ | |_ __ ___  
% | '_ ` _ \| | '_ ` _ \ 
% | | | | | | | | | | | |
% |_| |_| |_|_|_| |_| |_|
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We begin with an example.

Imagine that we have a large collection of human cells.  We will call each cell a `specimen.'  For each specimen we have:
%
\begin{enumerate}
    \item $\ell$, the sampling strategy used to obtain the specimen.  We will also call $\ell$ the specimen ``subpopulation,'' for each strategy should capture a different subpopulation of the overall set of specimens.  For example, we could have a few strategies based on selecting cells of particular sizes, a few strategies based on selecting cells from particular places in the body, and a few strategies based on based on selecting cells with particular proteomic markers.  Each additional sampling strategy shows us a different subpopulation of the cells and strengthens what kinds of inferences are possible. 
    \item $X$, the cell's `transcriptomic type,' as measured by looking at the transcriptomic activity in the cell.  
    \item $Y$, the cell's `morphological type,' as measured by looking at an image of the cell.
\end{enumerate}
%
Our task is to reconcile the two different notions of `cell-type' indicated by $X$ and $Y$, by constructing what we call a `calibration.'  This calibration should tell us what transcriptomic types are associated with what morphological types.  For example, we might like to be able to say ``This cell has transcriptomic type 3, therefore it probably has morphological type F.''  We can make math of this idea by defining a calibration as a conditional probability $q^*(y|x,\ell)$. This object indicate the likelihood that the morphological type is $y$ given that the transcriptomic type is $x$ and the specimen was sampled with strategy $\ell$.  In these terms, our task is to learn the calibration $q^*$.

Learning $q^*$ is hard when ``joint measurement'' is impossible or impractical.  By this we mean that for any given specimen we can either observe $\ell, X$ or $\ell,Y$.  We can never observe $\ell,X,Y$ for any specimen.  This makes it much more difficult to estimate the calibration; the fundamental difficulty is sketched in Figure \ref{fig:bigpicture}.  To make calibration possible without joint measurement, we need an assumption additional assumption.  The Markov Link Method supplies this assumption in the form of a conditional independence assumption.  This makes it possible to learn about the calibration.

To make it possible to learn $q^*$ even without joint measurement, the MLM makes the assumption that the calibration is the same for every sampling strategy:

\begin{center}
\fbox{\parbox[c]{4in}{
\vspace{.1in}
\begin{center}
\textbf{The Markov Link Method Assumption}\\$q^*(y|x,\ell)=q^*(y|x)$ for each $\ell$
\end{center}
\vspace{.1in}
}}
\end{center}

This assumption can be understood through the following thought experiment.  Let us say we are told that a cell has transcriptomic type $x$.  We could then make predictions about the morphological type $y$.  Now we are told new information: the cell was obtained via sampling strategy $\ell$.  Does this change our predictions about the morphological type?  If so, the MLM assumption is violated.  However, if $X$ is sufficiently informative then learning the sampling strategy will not change our predictions about the morphological type.  In this case, the MLM assumption holds.

If the MLM assumption holds, a frequentist point of view suggests the following procedure for estimating the calibration $q^*$ from data:

\figgy{
\hfill{}\figgygr{.9}{../images/qtildesamps}\hfill{}
\caption{{\bf Estimating the calibration and measuring our uncertainty}.  \label{fig:qtildesamps} On the left we visualize our best guess for true the calibration between two different tools for measuring cell-type.  Tool I uses one approach to obtain a fine-grained classification of cells, whereas tool II uses a different approach and only gives coarse type indication.  We can visualize our estimated calibration with a heatmap, formed by drawing a rectangle for each $x,y$ with a color indicating the magnitude of $\tilde q(y|x)$.  For example, examine the bright yellow square on the `Pvalb Gpr149' row and the `unknown' column.  This suggests that if a cell has the `Pvalb Gpr149' type according to tool I, it is extremely likely to be assigned the `unknown' type by tool II.  However, our best guess might not be correct.  On the right we show many plausible alternatives for the `Pvalb Gpr149' row of the calibration.  Each possibility is drawn from a 95\% confidence region for calibration.  Some of the alternatives suggest that some `Pvalb Gpr149' cells might not be classified as `unknown,' but might instead be correctly identified as being of the `Pvalb' type.  The data we have cannot tell us which of these alternatives is closest to the truth.  More data and more subpopulations would be necessary to reduce the variability within this confidence region.  However, we can have some confidence that `Pvalb Gpr149' cells are not being classified as `Meis2' or `Vip', insofar as none of the plausible calibrations give high probability to these tool II types.}
}

\begin{enumerate}
    \item Summarize our observed data in two matrices:
    \begin{align*}
    N^X_{\ell x} &= |\{i\leq n:\ \ell_i=\ell,X_i=x\}|\\
    N^Y_{\ell y} &= |\{j\leq m:\ \ell_j=\ell,Y_j=x\}|
    \end{align*}
    That is, $N^X_{\ell x}$ indicates the number of specimens from subpopulation $\ell$ with transcriptomic type $x$.  Likewise $N^Y_{\ell y}$ indicates the number of specimens from subpopulation $\ell$ which with morphological type $y$.

    \item Define a confidence region for calibration $q^*$.  To do this we incorporate a nuisance parameter: let $p^*(x|\ell)$ denote the probability that a cell sampled with strategy $\ell$ will have transcriptomic type $x$.  We form a joint simultaneous confidence region for all of $p^*,q^*$ using the following likelihood ratio statistic:
    \[
    L(p,q)=\frac{1}{\Omega_X}\sum_{\ell,x} N^X_{\ell x} \log \frac{\hat p(x|\ell)}{p(x|\ell)} + \frac{1}{\Omega_Y}\sum_{\ell,y} N^Y_{\ell y} \log \sum_x \frac{\hat h(y|\ell)}{p(x|\ell) q(y|x)}
    \]
    where $\hat p(y|x)=N^X_{\ell x} / \sum_{x'}N^X_{\ell x'}$ and $\hat h(y|x)=N^Y_{\ell y} / \sum_{y'}N^Y_{\ell y'}$ denote empirical conditional distributions.  We then define our confidence region as 
    \[
    R = \{(p,q):\ L(p,q) < k\}
    \]
    and choose $k$ to ensure any desired coverage.  For example, to produce a 95\% confidence region, we would like to choose $k$ so that the probability that $p^*,q^* \in R$ is at least 95\%.  We refer the reader to Appendix \ref{sec:confinterval} for details on choosing $k$.

    \item Determine whether the confidence region is empty.  If the region is empty then we can reject the MLM assumption hypothesis with a significance level of one minus the region coverage.  Assuming we fail to reject, we can continue:

    \item Sketch out the the confidence region, using examples.  In most nontrivial cases the region $R$ is high-dimensional and difficult to inspect directly.  We therefore produce examples from this region to try to help the user understand what values of $q$ are plausible.  In particular, we sample $Z_{x,y} \sim \mathrm{Exponential}(1)$ for each $x,y$ and solve the problem
    \[
    \max_{p,q \in R} \quad \sum_{x,y} Z_{x,y} \log q(y|x) \\
    \]
    This problem does not have a closed-form solution, but can generally be solved using an iterative procedure; we describe our method in detail in Appendix \ref{sec:optproblems}.  The result is a particular value of $q$ which lies within $R$.  In practice, we find these examples are almost always on the boundary of $R$.  We can then repeat this process by resampling $Z$.  By repeating this process many times, we can get a large collection of plausible values of $q$.  For any fixed value of $x$ we can visualize these samples by using a heatmap to plot $q(\cdot |x)$ for all the samples.  See Figure \ref{fig:qtildesamps} for an example.

    \item Produce a point estimate.  It is not generally possible to produce a consistent point estimate.  Indeed, there are many situations in which the confidence region $R$ does not concentrate even in the limit of infinite data.  Nonetheless, if a point estimate is desired, one may pick point estimates $\tilde p,\tilde q$ by solving the minimization problem $\min L(p,q)$.  We caution that this problem does not always have a unique solution, so the result may depend upon the minimization strategy used.  Precise details for our algorithm may be found in Appendix \ref{sec:optproblems}.
\end{enumerate}

A Bayesian point of view would suggest a different procedure; this point of view may provide a useful complementary perspective and is worthy of future research.  However, we caution that it is nontrivial to choose reasonable prior distributions for $p^*,q^*$.  Priors which may seem uninformative (such as the Dirichlet distribution) can actually carry significant and unanticipated weight, due to the high-dimensional nature of the problem.  We give an example in \ref{sec:bayesian}.

The frequentist confidence region allows us to avoid making any modeling assumptions.  However, we are then faced with the task of understanding a high-dimensional confidence region.  As described above, our solution is to look at examples from inside the region.  We emphasize that these examples are not drawn uniformly within the confidence region; instead the examples illustrate something more like the boundary of the region, sketching out what might be possible.  We concede that this method is somewhat ad-hoc and hope that future research may discover a superior method to visualize this high-dimensional confidence region.  For now, it appears to be a practical solution.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                 _           
%   _____  ____ _ _ __ ___  _ __ | | ___  ___ 
%  / _ \ \/ / _` | '_ ` _ \| '_ \| |/ _ \/ __|
% |  __/>  < (_| | | | | | | |_) | |  __/\__ \
%  \___/_/\_\__,_|_| |_| |_| .__/|_|\___||___/
%                          |_|               
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\label{sec:examples}

@@

\begin{itemize}
    \item Quality control for manufacturing.   One way to test the reliability of a part is to construct a machine that pushes the part until it breaks.  However, how can we test the reliability of the machine that performs the test?  In each test run there will be some variability induced by the machine itself, which induces a measurement error.  In practice, some kind of assumptions about part homogeneity are used to approximate this error (cf.\ \cite{de2005gauge}).  However, if we have two testing machines we can use the MLM to obtain a calibration between the machines, even though we can never test the same part with both machines.  This enables us to bound the overall measurement error.  In this case, $\ell$ might indicate the type of a part being tested, $X$ would indicate the reliability of a part as measured by one machine, and $Y$ would indicate the reliability of a part as measured by another machine.  If the error in machine $Y$ is not correlated to the part type $\ell$, then the MLM assumption certainly holds.  Even if the error is correlated, the MLM assumption may still hold.  For example, imagine that the $Y$ error is correlated with the absolute reliability of the part; this may pose no problem if that reliability is adequately measured by $X$.

    \item Combining knowledge across experimental modalities: morphology and transcriptomics.  There are different ways to think about the different types of cells in an organism.  A traditional approach is to classify cells based on what they look like (cf.\ \cite{steinman1973identification,bloomfield1982physiological}).  A more modern approach is to assay the cell's transcriptome (cf.\ \citep{tasic2017shared}).  Unfortunately, modern high-resolution cell photography and single-cell sequencing technologies are both destructive.  As a result, we can't always get both kinds of data for the same specimens.  For cells native to regions full of diverse cell-types, it is thus quite hard to grasp the correspondence between these different kinds of classification systems.  The result is two completely independent classifications of cells, one for each way of looking at the cell.  MLM allows us to estimate the relationship between those two classification systems, yielding a holistic understanding of the different types of cells.  In this case, $\ell$ might indicate some side information such as where in the body the cell was found, $X$ would indicate a detailed classification of the cell according to its transcriptomics, and $Y$ would indicate a coarser classification of a cell according to its morphology.  We expect that cell morphology is largely a function of cell transcriptomics.  Thus, as long as the $X$ measurement is sufficiently detailed, we expect that any correlations between $Y$ and $\ell$ would be explained by $X$.  That is, the MLM assumption holds.

@@

    \item Text/image correspondence.  Automatic image captioning is an ongoing effort in machine learning (cf.\ \cite{srivastava2018survey}).  There are three types of data available to help develop such algorithms: text-only data, image-only data, and paired-text-and-image data.  Obviously the last kind is the most useful for automatic image captioning, but there is much less of it.  The Markov Link Method suggests one way to use the more plentiful text-only and image-only data.  We can first apply classic machine learning techniques to get coarse labels for both kinds of data.  Using this side-information to identify subpopulations, the MLM can then deduce a fine-grained correspondence between text and images by combining information from across all the subpopulations.  Here $\ell$ would indicate coarse labels such as ``cat'' or ``street scene.''  These labels could be derived from either images or text and can be trained in a supervised fashion.  $X$ would indicate the image and $Y$ would indicate a caption.  If the caption is largely determined by the picture $X$, the MLM assumption may hold.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%      _                 _       _   _                 
%  ___(_)_ __ ___  _   _| | __ _| |_(_) ___  _ __  ___ 
% / __| | '_ ` _ \| | | | |/ _` | __| |/ _ \| '_ \/ __|
% \__ \ | | | | | | |_| | | (_| | |_| | (_) | | | \__ \
% |___/_|_| |_| |_|\__,_|_|\__,_|\__|_|\___/|_| |_|___/
%                                                    
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Simulation results}

\label{sec:simulations}

We use simulations to test the capabilities of the Markov Link Method to estimate a calibration.  We will:

\begin{itemize}
    \item Pick parameters to be `ground truth': 
    \begin{itemize}
        \item A calibration, $q^*(y|x)$.  
        \item A tool-I measurement distribution, $p^*(x|\ell)$.  Recall that this indicates the probability that tool I measures $x$ for a specimen gathered with sampling strategy $\ell$. 
        \item A tool-II measurement distribution, $h^*(y|\ell)$.   This indicates the probability that tool II measures $y$ for a specimen gathered with sampling strategy $\ell$.  We assume the the MLM assumption holds, and so this can be calculated from $q^*,p^*$ by the formula $h^*(y|\ell)=\sum_x p^*(x|\ell)q(y|x)$.  
    \end{itemize}
    \item Select a number of samples to take per subpopulation, $s$, and produce a simulated dataset for each subpopulation $\ell$:
    \begin{align*}
    (N^X_{\ell 1},N^X_{\ell 2}\cdots N^X_{\ell \Omega_X}) \sim \mathrm{Multinomial}(s,p^*(\cdot | \ell))\\
    (N^Y_{\ell 1},N^Y_{\ell 2}\cdots N^Y_{\ell \Omega_Y}) \sim \mathrm{Multinomial}(s,h^*(\cdot | \ell))
    \end{align*}
    \item Apply the Markov Link Method to the generated samples to obtain a calibration estimate $\tilde q$.
    \item Compute the error of the estimate $\tilde q$, as measured by a kind of averaged total variation distance, namely
    \[
    \frac{1}{2\Omega_{\ell}} \sum_{\ell x} |\tilde q(y|x) - q^*(y|x)|
    \]
    This distance ranges between zero and one.  A distance of zero indicates that $\tilde q=q^*$, and a distance of one indicates that the two calibrations place mass on completely disjoint sets, i.e. $q(y|x)=0$ whenever $\tilde q(y|x)>0$ and vice-versa.     
\end{itemize}

\figgy{
\hfill{}\figgygrs{.4}{../images/simulationIandIIpq.png}{(a) `Ground truth' for $p^*,q^*$.}\hfill{}
\hfill{}\figgygrs{.5}{../images/simulationsIandII.png}{(b) MLM performance \\ in estimating $q^*$.}\hfill{}

\caption{{\bf The Markov Link Method accurately estimates the calibration, as long as there are enough subpopulations}.  \label{fig:simsIandII}  We pick parameters as `ground truth' for $p^*,q^*$; these are as shown in subfigure (a) as heatmaps.  We then create simulated datasets using these parameter values, use the Markov Link Method to produce its estimator, and see and measure the total variation error of the estimator.  This error ranges from zero to one, where zero indicates no error and one indicates complete disagreement.  In subfigure (b) we see that the MLM is able to estimate the ground truth with negligible error, as long as it has enough samples.  However, we also consider another simulation: we merge some of the subpopulations in the ground truth parameters so that they become indistingiushable. In this degraded setup, the MLM is unable to learn the true calibration no matter how many samples we have. }
}

For the first simulation, we will look at a case in which the MLM performs well.  We will assume there are twenty different samping strategies, and tools I and II both measure specimens as being in one of six categories.  Our particular choices of ground truth can be found in Figure \ref{fig:simsIandII}.    This figure also shows that the MLM is able to accurately estimate the ground truth calibration $q^*$, as long as there are enough samples.

We also used this simulation to test the coverage of the confidence region.  Recall that the `actual coverage' of a confidence interval indicates the probability that a confidence region contains the ground truth parameters.  This coverage depends upon the particular choice of ground truth.  The `nominal' coverage indicates the intended coverage of the interval, and may be more or less than the actual coverage.  We estimated the actual coverage for the ground truth we have chosen.  Taking 200 sample datasets with $s=80$ samples per subpopulation, we found that an MLM confidence region with 95\% nominal coverage has an actual coverage of about $99\%$.  In this sense our region is `conservative,' which was somewhat by design.  We refer the reader to Appendix \ref{sec:confinterval} for details.  We also investigated other values of $s$, finding that the model became less conservative with more samples.  For example, for datasets $s=800$ with the actual coverage appears to be approximately $97.5\%$.

In our second simulation, we consider a case where the MLM does not succeed in accurately estimating the calibration.   We use the same ground truth parameters as in first simulation, but we will merge groups of sampling strategies.  In particular, we will divide the sampling strategies into four groups, each with five members.   We will pretend that we cannot distinguish subpopulations within any given group.  In this case, Figure \ref{fig:simsIandII} shows that the Markov Link Method cannot accurately estimate the calibration, no matter how many samples we have.  This is due to a fundamental identifiability issue which is discussed in detail in Appendix \ref{sec:casestudies}.

Despite the fact that we could not get an accurate point estimate for the calibration in the second simulation, the confidence regions appear to remain on solid ground.  We took 200 simulated datasets with $s=400$ samples per subpopulation (1600 samples total) and computed MLM confidence regions with $95\%$ nominal coverage. We found an actual coverage of about $96.5\%$.

\figgy{
\hfill{}\figgygr{.7}{../images/simulationIII}\hfill{}
